Torrential Twitter? Measuring the Severity of Harassment when Canadian Female Politicians Tweet about Climate Change

### 2024-02-07 Outputs/Paper/Publication_version/paper_for_pub.qmd
overhaul + details

"F\*CK OFF WITH YOUR CLIMATE SCAM .. nazi traitor b!tch," reads a reply sent to a Tweet[^1] about climate change that originated on the account of Canada's Deputy Prime Minister (DPM) and Minister of Finance, the Honourable Chrystia Freeland. For DPM Freeland and women across all political parties and levels of government in Canada and other democratic nations, coping with online and offline harassment is a reality of serving as an elected official [@southern; @silencingwomen]. Known as "gendertrolling," this harassment is part of a global trend of violence against women [@mantilla2013; @wagner]. Gender-based harassment on Twitter is well documented in academic literature and inexplicably pro-environmental behaviors are often viewed as "feminine" [@Citron; @mediatingmisogyny; @Anshelm]. Female politicians who adopt strong policy stances are more likely to face online misogyny and harassment because it is not consistent with existing gender stereotypes and notions of power and poses a threat to male climate deniers [@courtemanche; @rheault; @degeus].

@@

To gain a deeper understanding of the intersection of climate change denialism and online gendered-harassment faced by Canadian female politicians, I measure the severity of harassment DPM and MP for University--Rosedale, Hon. Chrystia Freeland (from the Liberal Party), MP for Saanich--Gulf Islands, Elizabeth May (Co-Leader of the Green Party), and MP for Victoria, Laurel Collins (from the New Democratic Party) receive specifically when Tweeting about climate change and environmental policy on a seven-point scale and investigate the types of accounts behind the Tweets on a six-point scale. After manually coding Tweets to determine the severity of harassment and type of account, I then built a Bayesian hierarchical model to explore the relationship between the type of account, whether a user was verified through Twitter's subscription program, and which female politician the reply was directed toward to see whether certain combinations of these variables are more likely to lead to a reply which contains harassment. It is hypothesized that evidence of harassment will be detected in the replies and that Personal and Anonymous Twitter accounts belonging to right-wing and ideologically-driven users will send Tweets containing the most harassment to the female politicians.

Harassment or gendertrolling is defined as sexist or misogynistic remarks that target a person based on their gender or sexuality and have varying levels of severity [@wagner]. Severity of harassment includes:

1.  Positive

2.  Neutral

3.  Questioning authority

4.  Name-calling/Gender insults

5.  Vicious language

6.  Credible threats

7.  Hate speech

Type of account is defined as the dynamics of relationships between followers, followings, and lists and will be determined by analyzing the accounts' bio, profile image, and other characteristics [@singh; @uddin]. Types of accounts include:

1.  Personal

2.  Professional

3.  Bots

4.  Spammers

5.  Anonymous

6.  Suspended/deleted

My analysis reveals that 86% of all replies to Tweets about climate change and environmental policy that originated on the accounts of DPM Freeland, MP May, and MP Collins contain some form of harassment, ranging from Questioning authority to Hate speech. The remaining 14% of Tweets had either positive or neutral sentiments. Questioning the authority of female politicians, followed by Name-calling/Gender insults were the most common forms of harassment identified, accounting for over 80% of all Tweets. Harassing Tweets were most likely to be sent by Spammers or Anonymous accounts. My Bayesian hierarchical model suggests that there are no significant quantitative differences between the severity of harassment received by the three politicians, although qualitative differences, including word choices and tone were detected during the coding process. My [Discussion] section further explores these qualitative and quantitative differences, among other key findings.

In the remainder of this paper, I commence with the [Literature Review], followed by my [Theoretical Framework]. Next, I discuss my [Data and Methods]. Then, I highlight my [Results] and [Model], before providing further analysis and insights in the [Discussion] section. Finally, in the [Conclusion], I summarize my main findings.

@@

## Twitter & The Digital Public Sphere

@@

The past few decades have led to the structural transformation and fragmentation of the public sphere, moving from having a singular, "official" public sphere to a multitude of parallel smaller ones, called "subaltern counterpublics" and networked publics mediated by technological affordances [@brunshighfield; @fraser1991; @boyd2010]. Technology reorganized the flow of information in society, distinguishing networked publics from subaltern counterpublics and the "official" public. Discourse traditionally happened in mediated spaces, but the rise of social media allows it to happen everywhere, theoretically breaking down barriers of who has access to the walled garden of democratic discourse [@mediatingmisogyny; @trifiro]. Despite the potential for the formation of new publics online, including issue publics driven by shared policy interests such as climate change, the leading voices in networked publics largely continue to be white, male, Christian, and have free time (often dictated by patriarchal norms) to actively contribute to discussions [@bruns2023; @mantilla2013; @degeus]. Networked publics use trolling and harassment to undermine the authority of female politicians whose actions do not fit traditional gender roles and conceptions of power [@degeus; @mantilla2015; @megarry2014]. Consequently, women have and continue to be on the margins of the public sphere, both in physical spaces like the House of Commons and online, as underscored by an @ipu global survey which found that 82% of female politicians reported being subjected to psychological violence (including harassment, threats, and sexist comments). In Canada, specific examples of on-going marginalization can be seen through the lived experiences of women in politics, with only one female Prime Minister to date (the Rt. Hon. Kim Campbell) and 30% of federal MPs being female [@degeus; @hoc].

@@

## Harassment of Women in Politics

@@

## Data Collection

@@

## Data Cleaning

### Raw Data

@@

### Coded Data

After manually coding the 4,139 Tweets using my Codebook (see [Appendix]), I proceeded to the second round of cleaning. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analyzed and visualized my data using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

## Data Coding

I developed a comprehensive codebook building on previous research and a sample of Tweets. Severity of harassment is coded on a seven-point scale, with the seven categories being: Positive, Neutral, Questioning authority, Name-calling/Gender insults, Vicious language, Credible threats, and Hate speech. There is agreement among scholars from fields such as communication and media studies, women and gender studies, and political science that there are no universal definition or ways of measuring and characterizing severity of harassment, incivility, and hate speech [@mantilla2015; @tenove; @krook2020]. Through this study, I develop and test a new quantitative seven-point scale which contributes to on-going research by emphasizing everyday forms of harassment, such as microaggressions and questioning the authority of female politicians. My third category, Questioning authority, is a specific contribution, as it is a phenomenon recognized in the literature [@krook2020], but not typically included in coding schemes as a way to categorize harassment directed toward female politicians.

These more everyday forms of harassment are often missed during binary coding conducted by large language models (LLMs). My seven-point scale also combines the nuance from previous qualitative projects which employ content analysis and interviews, while recognizing the benefits of more statistical and quantitative approaches offered by computational political communication researchers.

To ensure intercoder reliability, an undergraduate research assistant coded a random selection of ten percent of the total number of Tweets in the sample. Intercoder reliability was calculated using the package `irr` [@irr] to check the extended percentage agreement among coders. Agreement for severity of harassment was 87.5%, while agreement for type of account was 76%. @lacy2015 suggest that results over 80% are acceptable and results between 67% and 80% are tentatively acceptable, with explanation (further discussed in [Limitations]).

## Data Limitations

@@

## Severity of Harassment

@@

## Type of Account

@@

<!-- length(unique(finalized_cleaned_tweets$username)) While some users sent multiple replies to either one, two, or all of the female politicians, over 75% of users only appeared once in the dataset. My type of account results are consistent with the findings of @theocharis and @ward2020, who reveal that "... uncivil behavior seems to be common among many frustrated citizens who take to Twitter to angrily criticize and insult politicians" [@theocharis, p. 9]. -->

@@

## Chrystia Freeland

@@

## Elizabeth May

@@

## Laurel Collins

@@

## Elizabeth May
The results for severity of harassment and type of accounts detected in MP May's sample of Tweets are largely consistent with my hypothesis, although there were slightly fewer Positive and Neutral Tweets than anticipated. 55% of Tweets in MP May's sample were coded as Questioning authority, while 28% fall under the Name-calling/Gender insults category (see @tbl-liz_severity). @tbl-liz_account emphasizes that combined, Spammers and Anonymous accounts total nearly 80% of all accounts in MP May's networked public -- a significant number compared to the 19% of Personal accounts. Given MP May's decades of climate advocacy and personal popularity, a higher number of Personal accounts was anticipated, but the number of Spammers and Anonymous accounts could be explained by the changing platform affordances.

@@

## Deliberative Democracy

@@

## Limitations

@@

In this paper, I examined Tweets explicitly discussing climate change and environmental policy sent to DPM Freeland, MP May, and MP Collins to measure the severity of harassment on a seven-point scale and type of accounts behind the Tweets. My results reveal that over 85% of Tweets sent to the three politicians contained some level of harassment. My model identified and predicted nearly no quantitative differences between Tweets sent to all three female politicians. Spammer and Anonymous accounts were most frequently associated with engaging with the female politicians on Twitter and being responsible for sending the more serve forms of harassment identified. The Tweets analyzed for this study generally matched the criteria of what constitutes healthy democratic discourse, with users with multiple, diverse viewpoints weighing in on the issue of climate change. It is crucial to ensure we continue studying and discussing the intersection of climate change denialism and misogyny, because if we cannot mediate the risks of gendertrolling at the highest echelons of power, then female politicians may self-censor and risk everyone's ability to engage in respectful discourse surrounding climate change [@mediatingmisogyny].

>>>>

"F\*CK OFF WITH YOUR CLIMATE SCAM... nazi traitor b!tch," reads a reply sent to a Tweet[^1] about climate change that originated on the account of Canada's Deputy Prime Minister (DPM) and Minister of Finance, the Honourable Chrystia Freeland. For DPM Freeland and women across all political parties and levels of government in Canada and other democratic nations, coping with online and offline harassment is a reality of serving as an elected official [@southern; @trifiro]. Known as "gendertrolling," this harassment is part of a global trend of violence against women [@mantilla2013; @wagner]. Gender-based harassment on Twitter is well documented in academic literature and, inexplicably, pro-environmental behaviors are often viewed as "feminine" [@Citron; @mediatingmisogyny; @Anshelm]. Female politicians who adopt strong policy stances are more likely to face online misogyny and harassment because it is not consistent with existing gender stereotypes and notions of power and poses a threat to male climate deniers [@courtemanche; @rheault; @degeus].

@@

To gain a deeper understanding of the intersection of climate change denialism and online gendered-harassment faced by Canadian female politicians, I collected and manually coded replies to Tweets from DPM and MP for University--Rosedale, Hon. Chrystia Freeland (from the Liberal Party), MP for Saanich--Gulf Islands, Elizabeth May (Co-Leader of the Green Party), and MP for Victoria, Laurel Collins (from the New Democratic Party) that specifically discuss climate change and environmental policy. Tweets were coded using a seven-point scale to measure the severity of harassment and a six-point scale to categorize the types of accounts behind the replies. I then built a Bayesian hierarchical model to explore the relationship between the type of account, whether a user was verified through Twitter's subscription program, and which female politician the reply was directed toward to see whether certain combinations of these variables are more likely to lead to a harassing reply.

My analysis reveals that 86% of all replies to Tweets about climate change and environmental policy that originated on the accounts of DPM Freeland, MP May, and MP Collins contain some form of harassment, ranging from Questioning authority to Hate speech. Questioning the authority of female politicians, followed by Name-calling/Gender insults were the most common forms of harassment identified, accounting for over 80% of all Tweets. Harassing Tweets were most likely to be sent by Spammers or Anonymous accounts. My Bayesian hierarchical model suggests that there are no significant quantitative differences between the severity of harassment received by the three politicians, although qualitative differences, including word choices and tone were detected during the coding process.

This study contributes to our knowledge of online harassment faced by female politicians by introducing a seven-point scale to measure the severity of harassment and a six-point scale to categorize the types of accounts behind the replies. The third category of harassment, questioning authority, is a specific contribution, as it is a phenomenon recognized in the literature [@krook2020], yet not typically included in coding schemes designed to measure harassment. Lastly, this study advances our understanding of the intersection between climate change denialism and online misogyny, revealing that the combination of the two areas is toxic for female politicians and democratic discourse.

@@

## Twitter & The Digital Public Sphere {.unnumbered}

@@

The past few decades have led to the structural transformation and fragmentation of the public sphere, moving from having a singular, "official" public sphere to a multitude of parallel smaller ones, called "subaltern counterpublics" and networked publics mediated by technological affordances [@brunshighfield; @fraser1991; @boyd2010]. Technology reorganized the flow of information in society, distinguishing networked publics from subaltern counterpublics and the "official" public. Discourse traditionally happened in mediated spaces, but the rise of social media allows it to happen everywhere, theoretically breaking down barriers of who has access to the walled garden of democratic discourse [@mediatingmisogyny; @trifiro]. Despite the potential for the formation of new publics online, including issue publics driven by shared policy interests such as climate change, the leading voices in networked publics largely continue to be white, male, Christian, and have free time (often dictated by patriarchal norms) to actively contribute to discussions [@bruns2023; @mantilla2013; @degeus; @tweetingclimate]. Networked publics use trolling and harassment to undermine the authority of female politicians whose actions do not fit traditional gender roles and conceptions of power [@degeus; @mantilla2015; @megarry2014]. Consequently, women have and continue to be on the margins of the public sphere, both in physical spaces like the House of Commons and online, as underscored by an @ipu global survey which found that 82% of female politicians reported being subjected to psychological violence (including harassment, threats, and sexist comments). In Canada, specific examples of on-going marginalization can be seen through the lived experiences of women in politics, with only one female Prime Minister to date (the Rt. Hon. Kim Campbell) and 30% of federal MPs being female [@degeus; @hoc].

@@

## Harassment of Women in Politics {.unnumbered}

@@

## Data Collection {.unnumbered}

@@

## Data Cleaning {.unnumbered}

### Raw Data {.unnumbered}

@@

### Coded Data {.unnumbered}

After manually coding the 4,139 Tweets using my Codebook, I proceeded to the second round of cleaning. I eliminated columns of data which were useful for determining the type of account but not relevant for the next steps of my analysis. For my severity of harassment data, I updated value 1 to "Positive," 2 to "Neutral," 3 to "Questioning authority," 4 to "Name-calling/Gender insults," 5 to "Vicious language," 6 to "Credible threats," and 7 to "Hate speech" based on my Codebook. Furthermore, for my type of account data, I changed value 1 to "Personal," 2 to "Professional," 3 to "Bots," 4 to "Spammers," 5 to "Anonymous," and 6 to "Suspended/deleted" based on my Codebook. I tested the cleaned data using `validate` [@validate], primarily validating variable types and ensuring the number of rows in the dataset matches the number of Tweets I coded. I then analyzed and visualized my data using functions from `knitr` [@rknitr], `kableExtra` [@rKableExtra], `Formattable` [@rFormattable], and `ggplot2` [@rGgplot2].

## Data Coding {.unnumbered}

I developed a comprehensive codebook building on previous research and a sample of Tweets. Severity of harassment is coded on a seven-point scale, with the seven categories being: Positive, Neutral, Questioning authority, Name-calling/Gender insults, Vicious language, Credible threats, and Hate speech. There is agreement among scholars from fields such as communication and media studies, women and gender studies, and political science that there are no universal definition or ways of measuring and characterizing severity of harassment, incivility, and hate speech [@mantilla2015; @tenove; @krook2020]. Through this study, I develop and test a new quantitative seven-point scale which contributes to on-going research by emphasizing everyday forms of harassment, such as microaggressions and questioning the authority of female politicians [@krook2020]. My third category, Questioning authority, is a specific contribution, as it is not typically included in coding schemes as a way to categorize harassment directed toward female politicians. These more everyday forms of harassment are often missed during binary coding conducted by large language models (LLMs). My seven-point scale also combines the nuance from previous qualitative projects which employ content analysis and interviews, while recognizing the benefits of more statistical and quantitative approaches offered by computational political communication researchers.

To ensure validity, an undergraduate research assistant coded a random selection of ten percent (413) of the total number of Tweets in the sample. Intercoder reliability was calculated using the package `irr` [@irr] to check the extended percentage agreement among coders. Acceptable levels of agreement were reached, with agreement for severity of harassment at 87.5% and type of account at 76% [@lacy2015].

## Data Limitations {.unnumbered}

@@

## Severity of Harassment {.unnumbered}

@@

## Type of Account {.unnumbered}

@@

@@

## Chrystia Freeland {.unnumbered}

@@

## Elizabeth May {.unnumbered}

@@

## Laurel Collins {.unnumbered}

@@

## Elizabeth May {.unnumbered}
The results for severity of harassment and type of accounts detected in MP May's sample of Tweets are largely consistent with my hypothesis, although there were slightly fewer Positive and Neutral Tweets than anticipated. 55% of Tweets in MP May's sample were coded as Questioning authority, while 28% fall under the Name-calling/Gender insults category (see @tbl-liz_severity). @tbl-liz_account emphasizes that combined, Spammers and Anonymous accounts total nearly 80% of all accounts in MP May's networked public -- a significant number compared to the 19% of Personal accounts. Given MP May's decades of climate advocacy and personal popularity, a higher number of Personal accounts was anticipated, but the number of Spammers and Anonymous accounts could be explained by the changing platform affordances.

@@

## Deliberative Democracy {.unnumbered}

@@

## Limitations {.unnumbered}

@@

This study examines replies to Tweets explicitly discussing climate change and environmental policy sent to DPM Freeland, MP May, and MP Collins. My results from manually coding and my Bayesian hierarchical model suggest that the three politicians are equally impacted by harassment, with 86% of Tweets containing some form of harassment. Questioning authority and name-calling/gender insults were the most identified forms of harassment, while Spammers and Anonymous accounts were most frequently associated with sending harassing replies. This study contributes to our understanding of the online harassment of female politicians by developing a seven-point scale to measure more nuanced forms of harassment and a six-point scale to categorize the types of accounts behind the replies

The Tweets analyzed for this study generally match the criteria of what constitutes healthy democratic discourse, with users with multiple, diverse viewpoints weighing in on the issue of climate change. It is crucial to ensure we continue studying and discussing the intersection of climate change denialism and misogyny, because if we cannot mediate the risks of gendertrolling at the highest echelons of power, then female politicians may self-censor and risk everyone's ability to engage in respectful discourse surrounding climate change [@mediatingmisogyny].
